\chapter{Use \soft{git} for your project}

In this chapter we will see how to start using \soft{git} within :\\
- a new project \\
- an existing project already tracked by \soft{git} \\
- an existing project tracked by \soft{svn} \\

\section{Creation of a new \soft{git} repository}

In a terminal, go to the directory where you want to start the version control:
\begin{lstlisting}
  cd \$PATH/myDirectory
\end{lstlisting} %$
The directory may be empty or already contain files, it doesn't matter. Initialise \soft{git} in the directory from the terminal by entering:
\begin{lstlisting}
  git init
\end{lstlisting}
\soft{git} is now enabled for your repository.
You need to indicate your name and email that will be associated to the commits made in this repository by entering in the terminal:
\begin{lstlisting}
  git config user.name "My Name"
  git config user.email myName@edf.fr
\end{lstlisting}
You can also configure \soft{git} to use the text editor of your choice for making commits.
Otherwise the default \soft{git} interface will be used.
If you want to set your text editor (\soft{vi} or \soft{emacs} are recommended for this) type:
\begin{lstlisting}
  git config core.editor vim
\end{lstlisting}

Then, you need to tell \soft{git} which files you want to track.
Imagine that at the beginning of the project \soft{myDirectory} only contains the files \file{README.txt} and \file{CODINGRULES.txt}. To track changes in these files type:
\begin{lstlisting}
  git add README.txt CODINGRULES.txt
\end{lstlisting}

\important{Important}:
If you don't tell \soft{git} to add a file to the repository it won't track it.
Each time you create a new file in the directory, if you want \soft{git} to track it you need to type \important{git add newFile.ext}.

To have \soft{git} create a first version of your project type:
\begin{lstlisting}
  git commit
\end{lstlisting}
\soft{git} then pops up a text editor window and asks you to enter a \important{commit message}. For example, type:
\begin{lstlisting}
  First commit
\end{lstlisting}
then save (CTRL-X+CTRL-S in \soft{emacs}) and close the text editor.
\soft{git} then tells you a new commit has been made on the branch \soft{master} (we'll see later what branches are):
\begin{lstlisting}
 [master (root-commit) 11a6c58] First commit
 0 files changed
 create mode 100644 CODINGRULES.txt
 create mode 100644 README.txt

\end{lstlisting}


\important{Important: the \soft{git} repository should contain all the necessary files for your code to compile and execute: source files, makefile, case files, documentation, README.}


\section{Cloning an existing \soft{git} repository}

To clone an existing \soft{git} repository, type:
\begin{lstlisting}
  git clone name@domain:\$PATH/centralGitRepository
\end{lstlisting}%$
To clone from a \soft{github} repository this would look like:
\begin{lstlisting}
  git clone http://github.com/userName/myDirectory.git
\end{lstlisting}

You still need to configure \soft{git} by setting the user name and email as above, and choosing a text editor for the commits.

From inside EDF this is a bit more complicated due to the proxy.
Just create a repository, for example myRepository, then init git and configure the user and editor.
Then, open the file \soft{.git/config} with a text editor and add to it the following lines:
\begin{lstlisting}
  [remote "origin"]
    fetch = +refs/heads/*:refs/remotes/origin/*
    url = http://userName:pswd@github.com/repoUserName/repoName.git
  [http]
    proxy = http://proxypac.edf.fr:3128
  [https]
    proxy = https://proxypac.edf.fr:3128
\end{lstlisting}
where \important{userName} is your \soft{github} user name (you have to be registered as a contributor to the project by the project administrator).
\important{pswd} is your \soft{github} password (special characters must be ``translated'', a list is available at \url{http://meyerweb.com/eric/tools/dencoder/}).
\important{repoUserName} is the \soft{github} user name of the project administrator and \important{repoName} is the name of the repository.

Then, you can enter in the terminal:
\begin{lstlisting}
  git pull
\end{lstlisting}
and you'll have the remote repository as if you had cloned it.

\section{Cloning an existing \soft{svn} repository}

To clone an existing \soft{svn} repository, type:
\begin{lstlisting}
  git svn clone http://address/centralSvnRepository
\end{lstlisting}%$
To clone from the opentelemac repository this would look like:
\begin{lstlisting}
  git svn clone http://svn.opentelemac.org/svn/opentelemac/trunk
\end{lstlisting}
If the cloning is interrupted (due for instance to the EDF proxy limitations), to finish procedure, you need to run the following command from your repository:
\begin{lstlisting}
  git svn fetch
\end{lstlisting}
The clone thus created only contains a specific branch of the central \soft{svn} repository, tagged as \important{remotes/git-svn}.

If you want to add the \soft{svn} branch \verb"myBranch" you should edit: \verb"my_repository/.git/config" as following:
\begin{lstlisting}
[svn-remote "svn"]
	url = http://svn.opentelemac.org/svn/opentelemac/
	fetch = trunk:refs/remotes/git-svn
	fetch = branches/myBranch:refs/remotes/myBranch
\end{lstlisting}

and run the command:
\begin{lstlisting}
  git svn fetch
\end{lstlisting}

You can verify that a new remote branch was added with the command:
\begin{lstlisting}
  git branch -r
\end{lstlisting}

and you can now create a local branch from the remote branch:
\begin{lstlisting}
  git branch myBranch-local myBranch
\end{lstlisting}

You can work in this git repository following the advice given in the chapter~\ref{sec:useGit}.
Chapter~\ref{sec:gitsvn} shows how to share your developments with the \soft{svn} repository.
You still need to configure \soft{git} by setting the user name and email as above, and choosing a text editor for the commits.

From inside EDF, before doing the clone it is necessary to configure \soft{svn} for the proxy : you need to modify the \verb"$HOME/.subversion/server"
file by adding the following lines under the line \verb"[global]":
\begin{lstlisting}[language=bash]
[global]
...
http-proxy-host = proxypac.edf.fr
http-proxy-port = 3128
http-proxy-username = NNI
http-proxy-password = SesamePassword
...
\end{lstlisting}
Where:
\begin{itemize}
\item \textbf{http-proxy-host} is the address of your proxy.
\item \textbf{http-proxy-port} is the port of your proxy.
\item \textbf{http-proxy-username} is the login for your proxy.
\item \textbf{http-proxy-password} is the password for your proxy.
\end{itemize}
